---
title: "P2.Data.Exploration"
author: "MYC"
date: "10/25/2021"
output: 
  html_document:
    theme: cosmo
---
Loading all the default settings and preliminary programs.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing and all dplyr related programs
library(here)      #to set paths
library(tidymodels)      # for the recipes package, along with the rest of tidymodels
library(skimr)           # for variable summaries

```
Path to Processed Data and loading of cleaned data
```{r load data}
data_location <- here::here("data","processed_data","P2.Processed.rds")
data<- readRDS(data_location)
glimpse(data)
```
Summary tables
```{r All Summary}
#making a summary table of proportions of Ct
data%>%
  count(Tool)%>%
  mutate(Proportions = n/sum(n)) #sample/sum of sample

data%>%
  skim()
```
## The Undetected Ct Values
Data Split out data for undetected
```{r Undetected}
#Create df of only samples without any Ct values
Undetected<-data%>%
  filter(is.na(Ct)) #filter for all NA's in Ct value-- this will mean not detected or error

skim(Undetected)

Undetected%>%
  filter(!grepl("*.CTRL", Tool))%>% #!grepl= removes anything ending with .CTRL in the Tool column
  group_by(Detector, Method, Input, Dilution)%>%
  count(Tool)%>%
  arrange(desc(n))%>%
  ggplot(aes(y=n,))+
  geom_col(position="dodge2", aes(x=Method, fill=Dilution ))+
  facet_grid(.~Detector)+
  labs(title="Number of Tool samples that were undetected after sampling and extradtion")
```
From the current dataset we see that at an input of 300 from tool B have a chance of being undetected.
The sample with most detection errors is the automated method recovering at the -3 log of solution for both probes.

## Back to the Data as a whole
```{r}
data%>%
  count(Method, Tool, Dilution, Input, sort=TRUE)

summary(data)
```

```{r}
ToolSummary<-data%>%
  select(Method, Detector, Tool, Input, Dilution, Ct)%>%
  filter(Method!="PCR", Tool%in%c("A","B","SD"))%>% #Remove PCR controls
  group_by(Detector, Method, Tool, Input)%>%
  summarise(n=n())%>%
  arrange(n,Detector)
print(ToolSummary)
```

```{r}
data%>%
  filter(Tool%in%c("SD", "NEG.CTRL","POS.CTRL"))%>%
  ggplot(aes(Dilution, Ct))+
  geom_line()+
  facet_wrap(~Tool)+
  scale_y_reverse()
```

SD Histogram between NTP and CDC
```{r}
data%>%
  filter(Tool%in%c("SD", "NEG.CTRL","POS.CTRL"))%>%
  arrange(Tool%in%c("NEG.CTRL", "SD", "POS.CTRL"),
          Input%in%c("300","1000","*"))%>%
  ggplot(aes(Dilution, Ct))+
  geom_jitter(aes(color=Input))+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
  geom_hline(yintercept=37, linetype="longdash", color = "blue")+ #General confirmed threshold
  facet_wrap(~Method)+
  scale_y_reverse() #reverse the Y axis for Ct graphics, as lower Ct values indicate presence
```


