---
title: "Asks"
author: "MYC"
date: "10/27/2021"
output: html_document
---
Loading all the default settings and preliminary programs.
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing and all dplyr related programs
library(here)      #to set paths
library(tidymodels)      # for the recipes package, along with the rest of tidymodels
library(skimr)           # for variable summaries

```
List of asks:

* A vs B, 
* CDC vs NTP
* NTP 300 vs NTP 1000

```{r}
data%>%
  group_by(Detector, Method, Tool, Input)%>%
  filter(Method!="PCR")%>%
  summarize(n=n(),
            Ct=mean(Ct))%>%
  mutate(tools=(n/2))
```

Path to Processed Data and loading of cleaned data
```{r load data}
data_location <- here::here("data","processed_data","P2.Processed.rds")
data<- readRDS(data_location)
glimpse(data)
```

# A vs B

Tools are used to sample different concntrations of seeded virus.
A recovered from -1 inoculated coupon area.
B recovered from -3 inoculated coupon area.
```{r}
data%>% 
  filter(Tool%in% c("A", "B"))%>%
  count(Method)

data%>% 
  filter(Tool%in% c("A", "B"), Input!="6400")%>%
  ggplot(aes(x=Tool, y= Ct))+
  geom_boxplot(aes(color=Method))+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
  geom_hline(yintercept=37, linetype="longdash", color = "blue")+ #General confirmed threshold
  facet_grid(Detector~Input)+
  scale_y_reverse()+
  labs(y="Ct Value",
       x=NULL,
       title="Comparing A and B tool recovery")
  
```
Tool A has an average Ct recovery higher than B. This makes sense as the surfaces A recovered from is 2 logs higher (-1) than that of B (-3). A log drop is ~3Ct values we can see from the graphic that it trends as predicted.
Suggests that 


# CDC vs NTP

Compare the different extraction methods Serial Dilutions
```{r}
data%>%
  mutate(Dilution=as.numeric(as.factor(Dilution)))%>%
  filter(Tool=="SD", Input=="300")%>%
  ggplot(aes(x=Dilution, y=Ct, color=Method))+
  scale_y_reverse()+
  geom_point()+
  geom_smooth(aes(fill=Method),method=lm, alpha=0.20, fullrange=TRUE)+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
#  geom_hline(yintercept=37, linetype="longdash", color = "blue")+ #General confirmed threshold
  facet_wrap(Detector~Tool)+
  labs(title = "Serial Dilution comparisons between CDC and NTP Methods, Input at 300.")

```
Serial dilution data between CDC and NTP.
Noting that all SD are in the confirmed realm of under 27 Ct Value
CDC method is better than NTP at -1 concentration of the SD.
the -2 shows little difference between method recovery Ct values.
NTP appears to remain better than CDC at -3 and -4 levels.
Accounting for the SE looks like both methods are the same until the -2 dilution.
Might be significant difference at the -4 dilution.


Compare between extraction methods at input 300
```{r}
data%>%
  mutate(Dilution=as.numeric(as.factor(Dilution)))%>%
  filter(Input=="300")%>%
  ggplot(aes(x=Dilution, y=Ct, color=Tool))+
  scale_y_reverse()+
  geom_jitter()+
  geom_smooth(method=lm)+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
  geom_hline(yintercept=37, linetype="longdash", color = "blue")+ #General confirmed threshold
  facet_wrap(Method~Detector)+
  labs(title="Method extraction comparisons with standard input")

```



Compare between extraction methods with additional inputs
```{r}
data%>%
  mutate(Dilution=as.numeric(as.factor(Dilution)))%>%
  filter(Input!="CTRL")%>%
  ggplot(aes(x=Dilution, y=Ct, shape=Tool, color=Input))+
  scale_y_reverse()+
  geom_jitter()+
  geom_smooth(method=lm)+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
  geom_hline(yintercept=37, linetype="longdash", color = "blue")+ #General confirmed threshold
  facet_wrap(Method~Detector)+
  labs(title="Method extraction comparisons with all inputs and tools")
```


# NTP Input variations

Serial Dilution
```{r}
data%>%
  mutate(Dilution=as.numeric(as.factor(Dilution)))%>%
  filter(Method=="NTP", Tool=="SD", Input!="CTRL", Input!="6800")%>%
  ggplot(aes(x=Dilution, y=Ct, color=Input, shape=Input))+
  scale_y_reverse()+
  geom_jitter()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
  facet_wrap(.~Detector)+
  labs(title="Comparing Input levels on the Serial Dilituions with for the NTP method.")
```
NTP SD is similar to CDC SD in that the Ct values remain under 27, the most "confirmed" presence of nCoV-2.
Interesting! Looks like 300 input is really the lowest and as the inputs increase they kind of converge together at the -3 dilution level. 
Interpretation:
better detection at high copy levels with lower input; 
better detection at low copy levels with more input; 

If we could determine how many copies are in the -3 dilution we may be able to better estimate the capacity of the particles.

NTP
```{r}
data%>%
  mutate(Dilution=as.numeric(as.factor(Dilution)))%>%
  filter(Method=="NTP", Input=="300"|Input=="1000")%>%
  ggplot(aes(x=Dilution, y=Ct, shape=Input, color=Tool))+
  scale_y_reverse()+
  geom_jitter()+
  geom_smooth(method=glm, fullrange=TRUE)+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
  geom_hline(yintercept=37, linetype="longdash", color = "blue")+ #General confirmed threshold
  facet_wrap(.~Detector)+
  labs(title="NTP extraction comparisons with all at 300 and 1000 input levels and all tools")
```


NTP
```{r}
data%>%
  mutate(Dilution=as.numeric(as.factor(Dilution)))%>%
  filter(Method=="NTP", Tool!="SD", Input>=6000, Input!="CTRL")%>%
  ggplot(aes(x=Dilution, y=Ct, color=Input))+
  scale_y_reverse()+
  geom_boxplot(aes(fill=Input))+
  geom_hline(yintercept=27, linetype="longdash", color = "orange")+ #CDC confirmed threshold
  facet_grid(Tool~Detector)+
  labs(title="NTP extraction comparisons Eluents input levels and all tools")
```
Negatives show no detection. Positive controls are at -1 and have smaller Ct value than the A tools which recovered -1 inoculations. Gold line represents the threshold of 27

```{r}
data%>%
  filter(Input>=6000, Input!="CTRL")
```

