---
title: "Data Processing and Saving"
author: "MYC"
date: "10/29/2021"
output: html_document
---
This script loads the 2021-10-23_All.Update.csv, a composite of reports, from `raw-data`.
The script further processes, cleans, splits, and saves the data as rds files in the `processed_data` folder.

Loading all the default settings and preliminary programs.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing and all dplyr related programs
library(here)      #to set paths
library(lubridate) # for date manipulation
library(broom)    #for tidying up if needed

```
####
__Eventually this will read in the combined data frame made from separate reports__

For now, this will import a self-maintained csv file to do initial analysis.

###

## Loading File
```{r Load All Data}
data_location <- here::here("data","raw_data","2021-10-23_All.Update.csv")
rawdata <- read.csv(data_location)
```
## Clean Data

### View Data
```{r view data}
dplyr::glimpse(rawdata)

head(rawdata)
tail(rawdata)
summary(rawdata)
```
## Wrangle Data
```{r wrangle data}
All.PCR<-rawdata%>%
  #select data cols that will be relevant for analysis
  select(Phase:Ct)%>%
  #change some of the data types into other data types
  mutate(Phase=as.factor(Phase), 
         Plate=as.factor(Plate),
         Method=as.factor(Method),
         Tool=as.factor(Tool),
         Input=as.numeric(Input),
         Dilution=as.numeric(Dilution),
         Detector=as.factor(Detector),
         Ct=as.numeric(Ct))%>%
  glimpse()
#leaving Ct as Character, for not
summary(All.PCR)
```
```{r All Data}
save_data_location <- here::here("data","processed_data","ALL.Data-20211029.rds")

saveRDS(All.PCR, file = save_data_location)
```

## Split Data and Save

### Phase 1 select Data (P1)
```{r P1 Select}
Phase1Data<-All.PCR%>%
  filter(Phase=="1")

summary(Phase1Data)
```
```{r P1 Save RDS}
save_data_location <- here::here("data","processed_data","P1-20211029.rds")

saveRDS(Phase1Data, file = save_data_location)
```

### Phase 2 select Data (P2)
```{r P2 Select}
Phase2Data<-All.PCR%>%
  filter(Phase=="2")

summary(Phase2Data)

```
```{r P2 Save RDS}

save_data_location <- here::here("data","processed_data","P2-20211029.rds")

saveRDS(Phase2Data, file = save_data_location)
```